---
title: "Apply_function"
author: "Brian J. Knaus"
date: "April 27, 2016"
output: html_document
---


One of the weaknesses of R is that loops can be relatively slow to execute.
The apply family of functions attempts to address this.
Use `?apply` or `?lapply` for examples of other flavors of the apply command.

*****

## Use of apply()

Create a test matrix.

```{r}
tmp <- matrix(rep(1:3, times=3), ncol=3)
tmp
```


'Apply' the function 'sum' over rows.

```{r}
apply(tmp, MARGIN=1, sum)
```


'Apply' the function 'sum' over rows.

```{r}
apply(tmp, MARGIN=2, sum)
```


*****

## Use of custom functions


If the operation we wish to apply to a data structure exists as an R function, we can call it from the apply command.
We can also define our own functions to apply over a data structure.


In practice, if we wanted to get averages over a matrix, there are existing functions that should be used.
Here we'll create our own as an example.


```{r}
myMean <- function(x){
  sum(x)/length(x)
}

apply(tmp, MARGIN=1, myMean)

```


Through defining our own functions we can extract all sorts of summaries from data in a fairly efficient manner.
