---
title: "VCF import and manipulation"
author: "Brian J. Knaus"
date: "`r Sys.Date()`"
output: html_document
---


The output from most common variant callers includes the VCF format.
Here we use the R package vcfR to import and manipulate VCF data.


## Availabilty

The package vcfR is available on CRAN and GitHub.

CRAN: [vcfR](http://cran.r-project.org/package=vcfR)

GitHub: [vcfR](https://github.com/knausb/vcfR)


## Data import

VCF format data (either gzipped or not) can be imported into R using the `read.vcfR()` function.


```{r, results='hide'}
library(vcfR)
tas <- read.vcfR("./data/TASSEL_GBS0077.vcf.gz")
```


## "show" method

The show method can be used to provide a quick summary of the VCF data.

```{r}
tas
```


## "head" method

```{r}
head(tas)
```


## VCF format data

```{r, eval=FALSE}
vignette("vcf_data", package='vcfR')
```



## Objects of class vcfR


**meta**


```{r}
tas@meta[1:4]
```


**fix**


```{r}
tas@fix[1:4,1:8]
```


**gt**


```{r}
tas@gt[1:4,1:3]
```


## extract.gt()


Because the data in the 'gt' section is not strictly tabular, we need a tool to parse the information we need.
The function `extract.gt()` is meant to perform this task.


```{r}
dp <- extract.gt(tas, element="DP", as.numeric = TRUE)
dp[1:4,1:8]
```


## Variant coverage


```{r, fig.width=12, fig.align='center'}
boxplot(dp[,1:20], las=3, col=2:8)
```



## heatmap.bp()


```{r, fig.align='center', fig.width=12, fig.height=12}
heatmap.bp(dp[1:1000,], rlabels=FALSE)
```


## apply()


See seperate document.


## Missingness


```{r, fig.align='center', fig.width=12}
miss <- apply(dp, MARGIN=2, function(x){sum(is.na(x))})

par(mar=c(10,4,4,2))
barplot(miss, las=3)
par(mar=c(5,4,4,2))
```



```{r, fig.align='center', fig.width=8}
miss <- miss/nrow(dp)
hist(miss, col=8, breaks=seq(from=0,to=1,by=0.1))
```



## Filter samples on missingness



```{r}
tas <- tas[,c(TRUE, miss < 0.4)]
tas
```



## Repeat for variants


```{r}
dp <- extract.gt(tas, element="DP", as.numeric = TRUE)
```


```{r, fig.align='center', fig.width=12, fig.height=12}
heatmap.bp(dp[1:1000,], rlabels = FALSE)
```


```{r, fig.align='center', fig.width=12}
miss <- apply(dp, MARGIN=1, function(x){sum(is.na(x))})
hist(miss, col=8, seq(0, 41, by=1))
```


```{r, fig.align='center', fig.width=12}
tas <- tas[miss < 4,]
tas
```


```{r}
dp <- extract.gt(tas, element="DP", as.numeric = TRUE)
```


```{r, fig.align='center', fig.width=12, fig.height=12}
heatmap.bp(dp[1:1000,], rlabels=FALSE)
```






## Output to file


```{r, eval=FALSE}
write.vcf(tas, 'TASSEL_GBS0077_filtered.vcf.gz')
```